### 2024-01-17 Paper/template/sample-sigconf-biblatex.tex
init paper

%%
%% This is file `sample-sigconf-biblatex.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf-biblatex')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf-biblatex.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%
%%
\documentclass[sigconf,natbib=false]{acmart}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
	\providecommand\BibTeX{{%
			Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmlicensed}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{XXXXXXX.XXXXXXX}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references, obtained by selecting the acmnumeric BibLaTeX style.
%% The acmauthoryear BibLaTeX style switches to the "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the acmauthoryear style of
%% citations and references.
%%
%% Bibliography style
\RequirePackage[
datamodel=acmdatamodel,
style=acmnumeric,
]{biblatex}

%% Declare bibliography sources (one \addbibresource command per source)
\addbibresource{software.bib}
\addbibresource{sample-base.bib}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

	%%
	%% The "title" command has an optional parameter,
	%% allowing the author to define a "short title" to be used in page headers.
	\title{The Name of the Title Is Hope}

	%%
	%% The "author" command and its associated commands are used to define
	%% the authors and their affiliations.
	%% Of note is the shared affiliation of the first two authors, and the
	%% "authornote" and "authornotemark" commands
	%% used to denote shared contribution to the research.
	\author{Ben Trovato}
	\authornote{Both authors contributed equally to this research.}
	\email{trovato@corporation.com}
	\orcid{1234-5678-9012}
	\author{G.K.M. Tobin}
	\authornotemark[1]
	\email{webmaster@marysville-ohio.com}
	\affiliation{%
		\institution{Institute for Clarity in Documentation}
		\streetaddress{P.O. Box 1212}
		\city{Dublin}
		\state{Ohio}
		\country{USA}
		\postcode{43017-6221}
	}

	\author{Lars Th{\o}rv{\"a}ld}
	\affiliation{%
		\institution{The Th{\o}rv{\"a}ld Group}
		\streetaddress{1 Th{\o}rv{\"a}ld Circle}
		\city{Hekla}
		\country{Iceland}}
	\email{larst@affiliation.org}

	\author{Valerie B\'eranger}
	\affiliation{%
		\institution{Inria Paris-Rocquencourt}
		\city{Rocquencourt}
		\country{France}
	}

	\author{Aparna Patel}
	\affiliation{%
		\institution{Rajiv Gandhi University}
		\streetaddress{Rono-Hills}
		\city{Doimukh}
		\state{Arunachal Pradesh}
		\country{India}}

	\author{Huifen Chan}
	\affiliation{%
		\institution{Tsinghua University}
		\streetaddress{30 Shuangqing Rd}
		\city{Haidian Qu}
		\state{Beijing Shi}
		\country{China}}

	\author{Charles Palmer}
	\affiliation{%
		\institution{Palmer Research Laboratories}
		\streetaddress{8600 Datapoint Drive}
		\city{San Antonio}
		\state{Texas}
		\country{USA}
		\postcode{78229}}
	\email{cpalmer@prl.com}

	\author{John Smith}
	\affiliation{%
		\institution{The Th{\o}rv{\"a}ld Group}
		\streetaddress{1 Th{\o}rv{\"a}ld Circle}
		\city{Hekla}
		\country{Iceland}}
	\email{jsmith@affiliation.org}

	\author{Julius P. Kumquat}
	\affiliation{%
		\institution{The Kumquat Consortium}
		\city{New York}
		\country{USA}}
	\email{jpkumquat@consortium.net}

	%%
	%% By default, the full list of authors will be used in the page
	%% headers. Often, this list is too long, and will overlap
	%% other information printed in the page headers. This command allows
	%% the author to define a more concise list
	%% of authors' names for this purpose.
	\renewcommand{\shortauthors}{Trovato et al.}

	%%
	%% The abstract is a short summary of the work to be presented in the
	%% article.
	\begin{abstract}
		A clear and well-documented \LaTeX\ document is presented as an
		article formatted for publication by ACM in a conference proceedings
		or journal publication. Based on the ``acmart'' document class, this
		article presents and explains many of the common variations, as well
		as many of the formatting elements an author may use in the
		preparation of the documentation of their work.
	\end{abstract}

	%%
	%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
	%% Please copy and paste the code instead of the example below.
	%%
	\begin{CCSXML}
		<ccs2012>
		<concept>
		<concept_id>00000000.0000000.0000000</concept_id>
		<concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
		<concept_significance>500</concept_significance>
		</concept>
		<concept>
		<concept_id>00000000.00000000.00000000</concept_id>
		<concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
		<concept_significance>300</concept_significance>
		</concept>
		<concept>
		<concept_id>00000000.00000000.00000000</concept_id>
		<concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
		<concept_significance>100</concept_significance>
		</concept>
		<concept>
		<concept_id>00000000.00000000.00000000</concept_id>
		<concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
		<concept_significance>100</concept_significance>
		</concept>
		</ccs2012>
	\end{CCSXML}

	\ccsdesc[500]{Do Not Use This Code~Generate the Correct Terms for Your Paper}
	\ccsdesc[300]{Do Not Use This Code~Generate the Correct Terms for Your Paper}
	\ccsdesc{Do Not Use This Code~Generate the Correct Terms for Your Paper}
	\ccsdesc[100]{Do Not Use This Code~Generate the Correct Terms for Your Paper}

	%%
	%% Keywords. The author(s) should pick words that accurately describe
	%% the work being presented. Separate the keywords with commas.
	\keywords{Do, Not, Us, This, Code, Put, the, Correct, Terms, for,
		Your, Paper}

	\received{20 February 2007}
	\received[revised]{12 March 2009}
	\received[accepted]{5 June 2009}

	%%
	%% This command processes the author and affiliation and title
	%% information and builds the first part of the formatted document.
	\maketitle

	\section{Introduction}
	ACM's consolidated article template, introduced in 2017, provides a
	consistent \LaTeX\ style for use across ACM publications, and
	incorporates accessibility and metadata-extraction functionality
	necessary for future Digital Library endeavors. Numerous ACM and
	SIG-specific \LaTeX\ templates have been examined, and their unique
	features incorporated into this single new template.

	If you are new to publishing with ACM, this document is a valuable
	guide to the process of preparing your work for publication. If you
	have published with ACM before, this document provides insight and
	instruction into more recent changes to the article template.

	The ``\verb|acmart|'' document class can be used to prepare articles
	for any ACM publication --- conference or journal, and for any stage
	of publication, from review to final ``camera-ready'' copy, to the
	author's own version, with {\itshape very} few changes to the source.

	\section{Template Overview}
	As noted in the introduction, the ``\verb|acmart|'' document class can
	be used to prepare many different kinds of documentation --- a
	double-anonymous initial submission of a full-length technical paper, a
	two-page SIGGRAPH Emerging Technologies abstract, a ``camera-ready''
	journal article, a SIGCHI Extended Abstract, and more --- all by
	selecting the appropriate {\itshape template style} and {\itshape
		template parameters}.

	This document will explain the major features of the document
	class. For further information, the {\itshape \LaTeX\ User's Guide} is
	available from
	\url{https://www.acm.org/publications/proceedings-template}.

	\subsection{Template Styles}

	The primary parameter given to the ``\verb|acmart|'' document class is
	the {\itshape template style} which corresponds to the kind of publication
	or SIG publishing the work. This parameter is enclosed in square
	brackets and is a part of the {\verb|documentclass|} command:
	\begin{verbatim}
		\documentclass[STYLE]{acmart}
	\end{verbatim}

	Journals use one of three template styles. All but three ACM journals
	use the {\verb|acmsmall|} template style:
	\begin{itemize}
		\item {\texttt{acmsmall}}: The default journal template style.
		\item {\texttt{acmlarge}}: Used by JOCCH and TAP.
		\item {\texttt{acmtog}}: Used by TOG.
	\end{itemize}

	The majority of conference proceedings documentation will use the {\verb|acmconf|} template style.
	\begin{itemize}
		\item {\texttt{sigconf}}: The default proceedings template style.
		\item{\texttt{sigchi}}: Used for SIGCHI conference articles.
		\item{\texttt{sigplan}}: Used for SIGPLAN conference articles.
	\end{itemize}

	\subsection{Template Parameters}

	In addition to specifying the {\itshape template style} to be used in
	formatting your work, there are a number of {\itshape template parameters}
	which modify some part of the applied template style. A complete list
	of these parameters can be found in the {\itshape \LaTeX\ User's Guide.}

	Frequently-used parameters, or combinations of parameters, include:
	\begin{itemize}
		\item {\texttt{anonymous,review}}: Suitable for a ``double-anonymous''
		conference submission. Anonymizes the work and includes line
		numbers. Use with the \texttt{\acmSubmissionID} command to print the
		submission's unique ID on each page of the work.
		\item{\texttt{authorversion}}: Produces a version of the work suitable
		for posting by the author.
		\item{\texttt{screen}}: Produces colored hyperlinks.
	\end{itemize}

	This document uses the following string as the first command in the
	source file:
	\begin{verbatim}
	\end{verbatim}

	\section{Modifications}

	Modifying the template --- including but not limited to: adjusting
	margins, typeface sizes, line spacing, paragraph and list definitions,
	and the use of the \verb|\vspace| command to manually adjust the
	vertical spacing between elements of your work --- is not allowed.

	{\bfseries Your document will be returned to you for revision if
		modifications are discovered.}

	\section{Typefaces}

	The ``\verb|acmart|'' document class requires the use of the
	``Libertine'' typeface family. Your \TeX\ installation should include
	this set of packages. Please do not substitute other typefaces. The
	``\verb|lmodern|'' and ``\verb|ltimes|'' packages should not be used,
	as they will override the built-in typeface families.

	\section{Title Information}

	The title of your work should use capital letters appropriately -
	\url{https://capitalizemytitle.com/} has useful rules for
	capitalization. Use the {\verb|title|} command to define the title of
	your work. If your work has a subtitle, define it with the
	{\verb|subtitle|} command.  Do not insert line breaks in your title.

	If your title is lengthy, you must define a short version to be used
	in the page headers, to prevent overlapping text. The \verb|title|
	command has a ``short title'' parameter:
	\begin{verbatim}
		\title[short title]{full title}
	\end{verbatim}

	\section{Authors and Affiliations}

	Each author must be defined separately for accurate metadata
	identification.  As an exception, multiple authors may share one
	affiliation. Authors' names should not be abbreviated; use full first
	names wherever possible. Include authors' e-mail addresses whenever
	possible.

	Grouping authors' names or e-mail addresses, or providing an ``e-mail
	alias,'' as shown below, is not acceptable:
	\begin{verbatim}
		\author{Brooke Aster, David Mehldau}
		\email{dave,judy,steve@university.edu}
		\email{firstname.lastname@phillips.org}
	\end{verbatim}

	The \verb|authornote| and \verb|authornotemark| commands allow a note
	to apply to multiple authors --- for example, if the first two authors
	of an article contributed equally to the work.

	If your author list is lengthy, you must define a shortened version of
	the list of authors to be used in the page headers, to prevent
	overlapping text. The following command should be placed just after
	the last \verb|\author{}| definition:
	\begin{verbatim}
		\renewcommand{\shortauthors}{McCartney, et al.}
	\end{verbatim}
	Omitting this command will force the use of a concatenated list of all
	of the authors' names, which may result in overlapping text in the
	page headers.

	The article template's documentation, available at
	\url{https://www.acm.org/publications/proceedings-template}, has a
	complete explanation of these commands and tips for their effective
	use.

	Note that authors' addresses are mandatory for journal articles.

	\section{Rights Information}

	Authors of any work published by ACM will need to complete a rights
	form. Depending on the kind of work, and the rights management choice
	made by the author, this may be copyright transfer, permission,
	license, or an OA (open access) agreement.

	Regardless of the rights management choice, the author will receive a
	copy of the completed rights form once it has been submitted. This
	form contains \LaTeX\ commands that must be copied into the source
	document. When the document source is compiled, these commands and
	their parameters add formatted text to several areas of the final
	document:
	\begin{itemize}
		\item the ``ACM Reference Format'' text on the first page.
		\item the ``rights management'' text on the first page.
		\item the conference information in the page header(s).
	\end{itemize}

	Rights information is unique to the work; if you are preparing several
	works for an event, make sure to use the correct set of commands with
	each of the works.

	The ACM Reference Format text is required for all articles over one
	page in length, and is optional for one-page articles (abstracts).

	\section{CCS Concepts and User-Defined Keywords}

	Two elements of the ``acmart'' document class provide powerful
	taxonomic tools for you to help readers find your work in an online
	search.

	The ACM Computing Classification System ---
	\url{https://www.acm.org/publications/class-2012} --- is a set of
	classifiers and concepts that describe the computing
	discipline. Authors can select entries from this classification
	system, via \url{https://dl.acm.org/ccs/ccs.cfm}, and generate the
	commands to be included in the \LaTeX\ source.

	User-defined keywords are a comma-separated list of words and phrases
	of the authors' choosing, providing a more flexible way of describing
	the research being presented.

	CCS concepts and user-defined keywords are required for for all
	articles over two pages in length, and are optional for one- and
	two-page articles (or abstracts).

	\section{Sectioning Commands}

	Your work should use standard \LaTeX\ sectioning commands:
	\verb|section|, \verb|subsection|, \verb|subsubsection|, and
	\verb|paragraph|. They should be numbered; do not remove the numbering
	from the commands.

	Simulating a sectioning command by setting the first word or words of
	a paragraph in boldface or italicized text is {\bfseries not allowed.}

	\section{Tables}

	The ``\verb|acmart|'' document class includes the ``\verb|booktabs|''
	package --- \url{https://ctan.org/pkg/booktabs} --- for preparing
	high-quality tables.

	Table captions are placed {\itshape above} the table.

	Because tables cannot be split across pages, the best placement for
	them is typically the top of the page nearest their initial cite.  To
	ensure this proper ``floating'' placement of tables, use the
	environment \textbf{table} to enclose the table's contents and the
	table caption.  The contents of the table itself must go in the
	\textbf{tabular} environment, to be aligned properly in rows and
	columns, with the desired horizontal and vertical rules.  Again,
	detailed instructions on \textbf{tabular} material are found in the
	\textit{\LaTeX\ User's Guide}.

	Immediately following this sentence is the point at which
	Table~\ref{tab:freq} is included in the input file; compare the
	placement of the table here with the table in the printed output of
	this document.

	\begin{table}
		\caption{Frequency of Special Characters}
		\label{tab:freq}
		\begin{tabular}{ccl}
			\toprule
			Non-English or Math&Frequency&Comments\\
			\midrule
			\O & 1 in 1,000& For Swedish names\\
			$\pi$ & 1 in 5& Common in math\\
			\$ & 4 in 5 & Used in business\\
			$\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
			\bottomrule
		\end{tabular}
	\end{table}

	To set a wider table, which takes up the whole width of the page's
	live area, use the environment \textbf{table*} to enclose the table's
	contents and the table caption.  As with a single-column table, this
	wide table will ``float'' to a location deemed more
	desirable. Immediately following this sentence is the point at which
	Table~\ref{tab:commands} is included in the input file; again, it is
	instructive to compare the placement of the table here with the table
	in the printed output of this document.

	\begin{table*}
		\caption{Some Typical Commands}
		\label{tab:commands}
		\begin{tabular}{ccl}
			\toprule
			Command &A Number & Comments\\
			\midrule
			\texttt{{\char'134}author} & 100& Author \\
			\texttt{{\char'134}table}& 300 & For tables\\
			\texttt{{\char'134}table*}& 400& For wider tables\\
			\bottomrule
		\end{tabular}
	\end{table*}

	Always use midrule to separate table header rows from data rows, and
	use it only for this purpose. This enables assistive technologies to
	recognise table headers and support their users in navigating tables
	more easily.

	\section{Math Equations}
	You may want to display math equations in three distinct styles:
	inline, numbered or non-numbered display.  Each of the three are
	discussed in the next sections.

	\subsection{Inline (In-text) Equations}
	A formula that appears in the running text is called an inline or
	in-text formula.  It is produced by the \textbf{math} environment,
	which can be invoked with the usual
	\texttt{{\char'134}begin\,\ldots{\char'134}end} construction or with
	the short form \texttt{\$\,\ldots\$}. You can use any of the symbols
	and structures, from $\alpha$ to $\omega$, available in
	\LaTeX~\cite{Lamport:LaTeX}; this section will simply show a few
	examples of in-text equations in context. Notice how this equation:
	\begin{math}
		\lim_{n\rightarrow \infty}x=0
	\end{math},
	set here in in-line math style, looks slightly different when
	set in display style.  (See next section).

	\subsection{Display Equations}
	A numbered display equation---one set off by vertical space from the
	text and centered horizontally---is produced by the \textbf{equation}
	environment. An unnumbered display equation is produced by the
	\textbf{displaymath} environment.

	Again, in either environment, you can use any of the symbols and
	structures available in \LaTeX\@; this section will just give a couple
	of examples of display equations in context.  First, consider the
	equation, shown as an inline equation above:
	\begin{equation}
		\lim_{n\rightarrow \infty}x=0
	\end{equation}
	Notice how it is formatted somewhat differently in
	the \textbf{displaymath}
	environment.  Now, we'll enter an unnumbered equation:
	\begin{displaymath}
		\sum_{i=0}^{\infty} x + 1
	\end{displaymath}
	and follow it with another numbered equation:
	\begin{equation}
		\sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f
	\end{equation}
	just to demonstrate \LaTeX's able handling of numbering.

	\section{Figures}

	The ``\verb|figure|'' environment should be used for figures. One or
	more images can be placed within a figure. If your figure contains
	third-party material, you must clearly identify it as such, as shown
	in the example below.
	\begin{figure}[h]
		\centering
		\includegraphics[width=\linewidth]{sample-franklin}
		\caption{1907 Franklin Model D roadster. Photograph by Harris \&
			Ewing, Inc. [Public domain], via Wikimedia
			Commons. (\url{https://goo.gl/VLCRBB}).}
		\Description{A woman and a girl in white dresses sit in an open car.}
	\end{figure}

	Your figures should contain a caption which describes the figure to
	the reader.

	Figure captions are placed {\itshape below} the figure.

	Every figure should also have a figure description unless it is purely
	decorative. These descriptions convey what’s in the image to someone
	who cannot see it. They are also used by search engine crawlers for
	indexing images, and when images cannot be loaded.

	A figure description must be unformatted plain text less than 2000
	characters long (including spaces).  {\bfseries Figure descriptions
		should not repeat the figure caption – their purpose is to capture
		important information that is not already provided in the caption or
		the main text of the paper.} For figures that convey important and
	complex new information, a short text description may not be
	adequate. More complex alternative descriptions can be placed in an
	appendix and referenced in a short figure description. For example,
	provide a data table capturing the information in a bar chart, or a
	structured list representing a graph.  For additional information
	regarding how best to write figure descriptions and why doing this is
	so important, please see
	\url{https://www.acm.org/publications/taps/describing-figures/}.

	\subsection{The ``Teaser Figure''}

	A ``teaser figure'' is an image, or set of images in one figure, that
	are placed after all author and affiliation information, and before
	the body of the article, spanning the page. If you wish to have such a
	figure in your article, place the command immediately before the
	\verb|\maketitle| command:
	\begin{verbatim}
		\begin{teaserfigure}
			\includegraphics[width=\textwidth]{sampleteaser}
			\caption{figure caption}
			\Description{figure description}
		\end{teaserfigure}
	\end{verbatim}

	\section{Citations and Bibliographies}

	The use of \BibTeX\ for the preparation and formatting of one's
	references is strongly recommended. Authors' names should be complete
	--- use full first names (``Donald E. Knuth'') not initials
	(``D. E. Knuth'') --- and the salient identifying features of a
	reference should be included: title, year, volume, number, pages,
	article DOI, etc.

	Using the BibLaTeX system, the bibliography is included in your source
	document with the following command, placed just before the \verb|\end{document}| command:
\begin{verbatim}
	\printbibliography
\end{verbatim}

The command \verb|\addbibresource{bibfile}| declares the \BibTeX\ file to use
in the {\bfseries preamble} (before the command
``\verb|\begin{document}|'') of your \LaTeX\ source
	where ``\verb|bibfile|'' is the name, \emph{with} the ``\verb|.bib|'' suffix.
	Notice that \verb|\addbibresource| takes only one argument: to declare multiple files,
	use multiple instances of the command.

	Citations and references are numbered by default. A small number of
	ACM publications have citations and references formatted in the
	``author year'' style; for these exceptions, please pass the option \verb|style=acmauthoryear|
	to the \verb|biblatex| package loaded in the {\bfseries preamble} (before the command
	``\verb|\begin{document}|'') of your \LaTeX\ source.

		Some examples.  A paginated journal article \cite{Abril07}, an
		enumerated journal article \cite{Cohen07}, a reference to an entire
		issue \cite{JCohen96}, a monograph (whole book) \cite{Kosiur01}, a
		monograph/whole book in a series (see 2a in spec. document)
		\cite{Harel79}, a divisible-book such as an anthology or compilation
		\cite{Editor00} followed by the same example, however we only output
		the series if the volume number is given \cite{Editor00a} (so
		Editor00a's series should NOT be present since it has no vol. no.),
		a chapter in a divisible book \cite{Spector90}, a chapter in a
		divisible book in a series \cite{Douglass98}, a multi-volume work as
		book \cite{Knuth97}, a couple of articles in a proceedings (of a
		conference, symposium, workshop for example) (paginated proceedings
		article) \cite{Andler79, Hagerup1993}, a proceedings article with
		all possible elements \cite{Smith10}, an example of an enumerated
		proceedings article \cite{VanGundy07}, an informally published work
		\cite{Harel78}, a couple of preprints \cite{Bornmann2019,
			AnzarootPBM14}, a doctoral dissertation \cite{Clarkson85}, a
		master's thesis: \cite{anisi03}, an online document / world wide web
		resource \cite{Thornburg01, Ablamowicz07, Poker06}, a video game
		(Case 1) \cite{Obama08} and (Case 2) \cite{Novak03} and \cite{Lee05}
		and (Case 3) a patent \cite{JoeScientist001}, work accepted for
		publication \cite{rous08}, 'YYYYb'-test for prolific author
		\cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might
		contain 'duplicate' DOI and URLs (some SIAM articles)
		\cite{Kirschmer:2010:AEI:1958016.1958018}. Boris / Barbara Beeton:
		multi-volume works as books \cite{MR781536} and \cite{MR781537}. A
		couple of citations with DOIs:
		\cite{2004:ITE:1009386.1010128,Kirschmer:2010:AEI:1958016.1958018}. Online
		citations: \cite{TUGInstmem, Thornburg01, CTANacmart}.
		Data Artifacts: \cite{UMassCitations}.
		Software project: ~\cite{cgal,delebecque:hal-02090402}. Software Version: ~\cite{gf-tag-sound-repo,}. Software Module: ~\cite{cgal:lp-gi-20a}. Code fragment: ~\cite{simplemapper}.

		\section{Acknowledgments}

		Identification of funding sources and other support, and thanks to
		individuals and groups that assisted in the research and the
		preparation of the work should be included in an acknowledgment
		section, which is placed just before the reference section in your
		document.

		This section has a special environment:
		\begin{verbatim}
			\begin{acks}
				...
			\end{acks}
		\end{verbatim}
		so that the information contained therein can be more easily collected
		during the article metadata extraction phase, and to ensure
		consistency in the spelling of the section heading.

		Authors should not prepare this section as a numbered or unnumbered {\verb|\section|}; please use the ``{\verb|acks|}'' environment.

		\section{Appendices}

		If your work needs an appendix, add it before the
		``\verb|\end{document}|'' command at the conclusion of your source
	document.

	Start the appendix with the ``\verb|appendix|'' command:
	\begin{verbatim}
		\appendix
	\end{verbatim}
	and note that in the appendix, sections are lettered, not
	numbered. This document has two appendices, demonstrating the section
	and subsection identification method.

	\section{Multi-language papers}

	Papers may be written in languages other than English or include
	titles, subtitles, keywords and abstracts in different languages (as a
	rule, a paper in a language other than English should include an
	English title and an English abstract).  Use \verb|language=...| for
	every language used in the paper.  The last language indicated is the
	main language of the paper.  For example, a French paper with
	additional titles and abstracts in English and German may start with
	the following command
	\begin{verbatim}
		\documentclass[sigconf, language=english, language=german,
		language=french]{acmart}
	\end{verbatim}

	The title, subtitle, keywords and abstract will be typeset in the main
	language of the paper.  The commands \verb|\translatedXXX|, \verb|XXX|
	begin title, subtitle and keywords, can be used to set these elements
	in the other languages.  The environment \verb|translatedabstract| is
	used to set the translation of the abstract.  These commands and
	environment have a mandatory first argument: the language of the
	second argument.  See \verb|sample-sigconf-i13n.tex| file for examples
	of their usage.

	\section{SIGCHI Extended Abstracts}

	The ``\verb|sigchi-a|'' template style (available only in \LaTeX\ and
	not in Word) produces a landscape-orientation formatted article, with
	a wide left margin. Three environments are available for use with the
	``\verb|sigchi-a|'' template style, and produce formatted output in
	the margin:
	\begin{description}
		\item[\texttt{sidebar}:]  Place formatted text in the margin.
		\item[\texttt{marginfigure}:] Place a figure in the margin.
		\item[\texttt{margintable}:] Place a table in the margin.
	\end{description}

	%%
	%% The acknowledgments section is defined using the "acks" environment
	%% (and NOT an unnumbered section). This ensures the proper
	%% identification of the section in the article metadata, and the
	%% consistent spelling of the heading.
	\begin{acks}
		To Robert, for the bagels and explaining CMYK and color spaces.
	\end{acks}

	%%
	%% Print the bibliography
	%%
	\printbibliography

	%%
	%% If your work has an appendix, this is the place to put it.
	\appendix

	\section{Research Methods}

	\subsection{Part One}

	Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi
	malesuada, quam in pulvinar varius, metus nunc fermentum urna, id
	sollicitudin purus odio sit amet enim. Aliquam ullamcorper eu ipsum
	vel mollis. Curabitur quis dictum nisl. Phasellus vel semper risus, et
	lacinia dolor. Integer ultricies commodo sem nec semper.

	\subsection{Part Two}

	Etiam commodo feugiat nisl pulvinar pellentesque. Etiam auctor sodales
	ligula, non varius nibh pulvinar semper. Suspendisse nec lectus non
	ipsum convallis congue hendrerit vitae sapien. Donec at laoreet
	eros. Vivamus non purus placerat, scelerisque diam eu, cursus
	ante. Etiam aliquam tortor auctor efficitur mattis.

	\section{Online Resources}

	Nam id fermentum dui. Suspendisse sagittis tortor a nulla mollis, in
	pulvinar ex pretium. Sed interdum orci quis metus euismod, et sagittis
	enim maximus. Vestibulum gravida massa ut felis suscipit
	congue. Quisque mattis elit a risus ultrices commodo venenatis eget
	dui. Etiam sagittis eleifend elementum.

	Nam interdum magna at lectus dignissim, ac dignissim lorem
	rhoncus. Maecenas eu arcu ac neque placerat aliquam. Nunc pulvinar
	massa et mattis lacinia.

\end{document}
\endinput
%%
%% End of file `sample-sigconf-biblatex.tex'.